---
title: "Plateau Hierarchical Effects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plateau Hierarchical Effects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The PaGAn package supports the implementation of a variation of the plateau
relationships outlined in [Brewer *et al.* (2016)](https://doi.org/10.1111/2041-210X.12609)
through its inclusion as a type of hierarchical effect.  The univariate version of
the plateau relationship can be described through the use of the following parameters:

+ $\beta_1$ (where $\beta_1 > 0$) - Describing the magnitude of the ascending
part of the plateau relationship
+ $\beta_2$ (where $\beta_2 > 0$) - Describing the magnitude of the descending
part of the plateau relationship
+ $\hat{x}$ - Describing the location of the hypothetical apex of the plateau
relationship on the x-axis
+ $\delta$ (where $\delta > 0$) - Describing the offset (on the $y$-axis of the
plateau peak from the hypothetical apex

The marginal response, $y$, then takes the following functional form:

$$
y \left( x \right) = \textrm{min} \left\{\left( x - \hat{x} \right) \beta_1, \left( \hat{x} - x \right) \beta_2, -\delta \right\}
$$

Plateau relationships can be ill-behaved in situations where there is little or no
data to inform the posterior distribution of the ascending or descending components
of the relationship.  The locations on the $x$-axis where the ascending or descending
components drive the relationship are controlled by the parameterisation of the
relationship and so this can result in poor mixing of MCMC samplers under certain
conditions.  To avoid this we enforce some constraints on the plateau parameters
such that, in situations where the ascending or descending components fall outside
the range of the data, we enforce a conservative bias towards symmetric
parameterisations.  This is done by defining an intermediate stochastic node,
$\mathbf{z}$, that has the following distribution:

$$
\mathbf{z} \sim \mathcal{N}_2 \left( \mathbf{0}, \frac{1}{\tau} \left[\begin{matrix}
1 & \rho \\
\rho & 1
\end{matrix}\right]  \right)
$$
and define

$$
\begin{align}
\beta_1 &= \left| z_1 \right| \\
\beta_2 &= \left| z_2 \right|
\end{align}
$$
where $\tau$ ($\tau > 0$) and $\rho$ ($0 \leq \rho < 1$) are hyperparameters that define the scale of the ascending and descending coefficients and the correlation between them respectively.  The Cholesky decomposition of the variance-covariance matrix, $\mathbf{L}$ is therefore

$$
\mathbf{L} = \frac{1}{\sqrt{\tau}} \left[ \begin{matrix}
1 & 0 \\
\rho & \sqrt{1 - \rho^2}
\end{matrix} \right]
$$
```{r setup}
# library(PaGAn)
```
