---
title: "Simple IPM Example"
author: "Joseph Chipperfield"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simple IPM Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Integral Projection Models in PaGAn

There are a number of fantastic packages for performing analysis of data using integral projection models (IPMs) in R such as [IPMpack](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210x.12001) and [impr](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13683).  PaGAn also has its own implementation of Integral Projection Models. The PaGAn implementation of IPM has been designed with a focus on flexibility and an ability to embed the outputs from these models into further analysis. Under the hood PaGAn uses Bayesian techniques (powered by [NIMBLE](https://www.tandfonline.com/doi/full/10.1080/10618600.2016.1172487)) which means that full posterior estimates of the coefficients in vital rate functions, kernels, and population-level summaries.

```{r libraries}
# This chunk will be removed once the package is completed - in the meantime it loads the required libraries for the analysis and imports the relevant source files from the project GitHub pages
library(nimble)
library(parallel)
library(coda)
library(DHARMa)
library(ggplot2)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
