---
title: "Installation of PaGAn Dependencies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation of PaGAn Dependencies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## Dependencies in PaGAn

Whilst the installation of the PaGAn package itself is a relatively simple
affair, it is important to note that some of the packages that PaGAn depends
upon require some extra installation steps to ensure that the entire
functionality is available for your system. It is worth running the following
steps in a fresh session of R (i.e. one that does not have a lot of packages
already loaded into the workspace).

## R Version

PaGAn has been written to take advantage of spatial analysis libraries that
underwent a significant reorganisation in the move from R version 4.2.x to
version 4.3.0. To ensure that the functionality runs as expected, then it is a
requirement that you have R version 4.3.0 or higher installed on your system.

## RTools (for Windows Users)

If you are running R on Windows then you will need to have an extra set of
compilation tools available so that model objects built in NIMBLE can be
compiled and run. These sets of tools are collectively know as 'Rtools' and you
must download and install the collection of Rtools that correspond to the
version of R that you are running. The Rtools that relates to version 4.3.x of R
is 
[Rtools43](https://cran.r-project.org/bin/windows/Rtools/rtools43/rtools.html).
Once Rtools is installed then you will be able to run code or install packages
that require compilation. You can check that Rtools is installed successfully
by running the following:

```{r checkRTools}
if(!requireNamespace("pkgbuild", quietly = TRUE)) {
  install.packages("pkgbuild")
}
pkgbuild::check_rtools(TRUE)
```

## Bioconductor Packages

One of PaGAn's dependencies, [INLA](https://www.r-inla.org/), has two suggested
packages, 'graph' and 'Rgraphviz', that are on Bioconductor. The following code
will install the bioconductor package manager ('BiocManager') if it is not
already installed and then use that package manager to install the relevant
dependencies:

```{r bioconductorDependencies}
if(!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
BiocManager::install(c("graph", "Rgraphviz"), dependencies = TRUE, lib = .libPaths()[1], update = TRUE, ask = FALSE)
```

## Installation of Direct Dependencies

Now all that is left is to install the other direct dependencies of PaGAn. Note
that INLA is not on [CRAN](https://cran.r-project.org/) and so its repository
must be added to the list of repositories to search when installing the
packages.

```{r directDependencies}
install.packages(c("INLA", "future"), repos = c(
  getOption("repos"),
  INLA = "https://inla.r-inla-download.org/R/stable"
), dependencies = TRUE)
```

## Installation of PaGAn

Finally, it is possible to install PaGAn directly from it's
[GitHub repository](https://github.com/joechip90/PaGAn):

```{r PaGAnInstall}
if(!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
devtools::install_github("joechip90/PaGAn", dependencies = TRUE)
```
